ScPy: A SuperCollider Extension for Performing Numerical Computation via Embedded Python

### 2016-08-27 doc/paper.tex
hello world and minor corrections

Although designing a domain-specific language (DSL)~\cite{van2000domain} for this task was briefly
considered, we eventually decided to use Python for a number of reasons. First, Python includes an
excellent C API which makes embedding it within other languages trivial in comparison to many other
options.  Second, Python's syntax is easily readable and concise --- it clearly adheres to our
goals. Finally, Python has a massive number of community built libraries for performant scientific
and numerical computing. With a few small exceptions, which are clearly specific to spectral data
such as the phase vocoder, these libraries contained every one of the fundamental operations we
could think of.

It may seem as though embedding an interpreted language within an interpreted language could offer
no performance advantage. There is certainly some performance hit when switching languages and
transferring data. However, the biggest advantage comes not strictly from performance but from the
diversity of operations which become available with highly optimized implementations. NumPy is used
internally by ScPy for handling data arrays, so the entire NumPy library of mathematical functions
comes at no cost.  Any other Python library can easily be imported as well. Limited experiments
have been done with the SciPy library. It would also be possible to use Theano to do processing on
the GPU for extra performance. Implementing an equivalent to these popular Python libraries in pure
SuperCollider is certainly possible, but it would require a very large amount of work and would see
little widespread adoption since SuperCollider is ultimately a niche language.

@@

>>>>

Although designing a domain-specific language (DSL)~\cite{van2000domain} for this task was briefly
considered, we eventually decided to use Python for a number of reasons. First, Python includes an
excellent C API which makes embedding it within other languages trivial in comparison to many other
options.  Second, Python's syntax is easily readable and concise --- it clearly adheres to our
goals. Finally, Python has a massive number of community built libraries for performant scientific
and numerical computing. With a few small exceptions, which are clearly specific to spectral data
(such as the phase vocoder), these libraries contained every one of the fundamental operations we
could think of.

It may seem as though embedding an interpreted language within an interpreted language could offer
no performance advantage. There is certainly some performance hit when switching languages and
transferring data. However, the biggest advantage comes not strictly from performance but from the
diversity of operations which become available with highly optimized implementations. NumPy is used
internally by ScPy for handling data arrays, so the entire NumPy library of mathematical functions
comes at no cost.  Any other Python library can easily be imported as well. Limited experiments
have been done with the SciPy library. It would also be possible to use Theano to do processing on
the GPU for extra performance. Implementing an equivalent to these popular Python libraries in pure
SuperCollider is certainly possible, but it would require a very large amount of work and would see
no adoption outside of the audio computing community.

@@

\begin{listing}[H]
    \begin{minted}[linenos=true]{SuperCollider}
(PyOnce("
    # comments in Python have different syntax!
    print('Hello from Python!')
    print('x is', x)
", (x: 62)))
    \end{minted}
    \caption{Hello world with ScPy.}
    \label{lst:helloworld}
\end{listing}
